# Cross-stiches

Ruijie Zhang

## Load Functions
Firstly, we will need 4 functions in the function.R file. (There are actually 5 functions in the R file, but change_resolution is used in make_pattern.) 
```{r}
set.seed(1005705300)
library(imager) 
library(tidyverse) 
library(tidymodels) 
library(sp) 
library(scales)
library(cowplot)


## dmc is installed by the following code. Delete the # sign.
#devtools::install_github("sharlagelfand/dmc")


################################ change resolution ################################

change_resolution <- function(image_df, x_size)
{
  ## change_resolution(image_df, x_size) subsamples an image to produce
  ## a lower resolution image. Any non-coordinate columns in the data
  ## frame are summarized with their most common value in the larger
  ## grid cell.
  ##
  ## Input:
  ## - image_df: A data frame in wide format. The x-coordinate column MUST
  ##             be named 'x' and the y-coordinate column MUST be named 'y'.
  ##             Further columns have no naming restrictions.
  ## - x_size:   The number of cells in the x-direction. The number of cells
  ##             in the vertical direction will be computed to maintain the 
  ##             perspective. There is no guarantee that the exact number
  ##             of cells in the x-direction is x_size
  ##
  ## Output:
  ## - A data frame with the same column names as image_df, but with fewer 
  ##   entries that corresponds to the reduced resolution image.
  ##
  ## Example:
  ##   library(imager)
  ##   library(dplyr)
  ##   fpath <- system.file('extdata/Leonardo_Birds.jpg',package='imager') 
  ##   im <- load.image(fpath)
  ##   im_dat<- as.data.frame(im,wide = "c") %>% rename(R = c.1, G = c.2, B = c.3) %>%
  ##            select(x,y,R,G,B)
  ##   agg_image <- change_resolution(im_dat, 50)
  
  if(!require(sp)) {
    stop("The sp packages must be installed. Run install.packages(\"sp\") and then try again.")
  }
  if(!require(dplyr)) {
    stop("The dplyr packages must be installed. Run install.packages(\"dplyr\") and then try again.")
  }
  
  sp_dat <- image_df 
  gridded(sp_dat) = ~x+y
  
  persp = (gridparameters(sp_dat)$cells.dim[2]/gridparameters(sp_dat)$cells.dim[1])
  y_size = floor(x_size*persp)
  orig_x_size = gridparameters(sp_dat)$cells.dim[1]
  orig_y_size = gridparameters(sp_dat)$cells.dim[2]
  
  x_res = ceiling(orig_x_size/x_size)
  y_res = ceiling(orig_y_size/y_size)
  
  gt = GridTopology(c(0.5,0.5), c(x_res, y_res),
                    c(floor(orig_x_size/x_res), floor(orig_y_size/y_res)))
  SG = SpatialGrid(gt)
  agg = aggregate(sp_dat, SG, function(x) names(which.max(table(x)))[1] )
  agg@grid@cellsize <- c(1,1)
  df <- agg %>% as.data.frame %>% rename(x = s1, y = s2)  %>% select(colnames(image_df))
  
  return(df)
  
}

################################ process_image ################################
process_image <- function(image_file_name,k_list){
  ## process_image(image_file_name,k_list) apply a several k-means clustering.
  ## to the image_file. 
  ##
  ## Input:
  ## - image_file_name: A PNG or JPEG image.
  ## - k-list: A vector that represents the number of cluster centers.
  ##
  ## Output:
  ## - A list of k-means clustering information and the clusters, 
  ##   their associated RGB values and their nearest DMC thread color.
  ##
  ## - Example:
  ## clusters <- process_image(my_image, c(2:10))
im <- imager::load.image(image_file)
tidy_dat <- as.data.frame(im, wide = "c") %>% rename(R = c.1, G = c.2, B = c.3)
dat <- select(tidy_dat,c(-x,-y))

kclusts <-
  tibble(k = k_list) %>%
  mutate(
    kclust = map(k, ~kmeans(x = dat , centers = .x, nstart=4)),
    glanced = map(kclust, glance),
  )

clusterings <-
  kclusts %>%
  unnest(cols = c(glanced)) %>%
  mutate(centers=map(kclust,tidy), 
         tidy_dat=map(kclust, ~augment(.x,tidy_dat) %>%
                                                   rename(cluster=.cluster)))

 for (k in k_list){
   center <- clusterings[clusterings$k==k,]$centers[[1]]
   center <- center %>%
     mutate(col=rgb(R,G,B)) %>%
     mutate(dmc=map(col, ~dmc(.x)))
   clusterings[clusterings$k==k,]$centers[[1]] <- center
   }
return(clusterings)
}

################################ scree plot ################################
scree_plot <- function(cluster_info){
  ## scree_plot produces and plots a scree plot.
  ##
  ## Input:
  ## - cluster_info: A data frame that contains a several cluster information. 
  ##
  ## Output:
  ## - Returns a scree plot which the y-axis is total within clusters SS and x-axis
  ##   is the number of clusters.
  ##
  ## - Example:
  ## scree-plot(clusterings)
p1 <- ggplot(cluster_info, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()
  
## ratio
nclust = length(cluster_info$k)
ratio = rep(NA, nclust-1)
for (kk in 2:nclust) {
ratio[kk-1] = cluster_info$tot.withinss[kk]/cluster_info$tot.withinss[kk-1]
  }
plot_data <- data.frame(k = cluster_info$k[2:nclust],ratio)
p2 <- ggplot(plot_data, aes(x=k, y = ratio)) + geom_line()
plot_grid(p1,p2)
}

################################ colour strip ################################
colour_strips <- function(cluster_info){
  ## colour_strip Produces colour strips with the DMC colour 
  ## closest to the cluster center colour.
  ##
  ## Input:
  ## - cluster_info: A data frame that contains a several cluster information. 
  ##
  ## Output:
  ## - colour strips with the DMC colour closest to the cluster center colour.
  ##
  ## - Example:
  ## colour_strips(cluster_info)
square <- function(x, label_size) { 
  ggplot()  + 
    coord_fixed(xlim=c(0,1), ylim = c(0,1)) + theme_void() + 
    theme(plot.background = element_rect(fill = x)) + 
    geom_text(aes(0.5,0.5),label = x , size = label_size)
}
strips <- list()
colour_ <- vector()
for( i in 1:nrow(cluster_info)){
  for(x in 1:cluster_info$k[i]){
    colour_[x] <- cluster_info$centers[[i]]$dmc[[x]]$hex
  }
  t <- tibble(colours = colour_,
              squares = purrr::map(colours, ~ square(.x, 2.5)))
  n_col = length(t$colours)
  rect_dat <- tibble(x1 = c(0:(n_col-1)), x2 = c(1:n_col), y1 = rep(0,n_col),
                     y2 =rep(1,n_col), colour = t$colours)
  strip <- rect_dat %>% ggplot() + coord_fixed() + 
    geom_rect(aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2, fill=colour), color="black") +
    geom_text(aes(x=x1+(x2-x1)/2, y=y1+(y2-y1)/2, label=colour), size=2.5) + 
    scale_fill_manual(values = rect_dat$colour)+ theme_void() + theme(legend.position = "none") 
  strips[[i]] <- strip
  }
plot_grid(plotlist = strips,
          labels=paste0("k=",c(cluster_info$k)),ncol=1)
}
################################  plot pattern ################################
make_pattern <- function(cluster_info,k,x_size,black_white=FALSE,background_color=NULL){
  ## plot_pattern plots the pattern of the given number of cluster center.
  ## 
  ##
  ## Input:
  ## - cluster_info: A data frame that contains a several cluster information. 
  ## - k: A numeric that restores the chosen cluster size.
  ## - x_size: A numeric that restores total number of possible stitches in the horizontal direction
  ## - black_white: (logical) Print the pattern in black and white (TRUE) or colour (FALSE)
  ##                The default is FALSE
  ## - background_colour: The colour of the background, which should not be stitched in the
  ##                      pattern. (Default is to not have a colour)
  ## Output:
  ## - Returns a cross-stitch pattern.
  ##
  ## - Example:
  ## make_pattern(cluster_info=clusters, k= 8,x_size=50 ,black_white = FALSE, background_color= NULL)
image_df <- cluster_info[cluster_info$k==k,]$tidy_dat[[1]]
lowres_df <- change_resolution(image_df,x_size)
cluster_dmc <- cluster_info[cluster_info$k==k,]$centers[[1]]


cluster_dmc$dmc_colour <- NA
for(i in 1:k){
  cluster_dmc$dmc_colour[i] <- cluster_dmc$dmc[[i]]$dmc
}

lowres_df$dmc_colour <- NA
lowres_df$name <- NA
for (i in 1:nrow(lowres_df)){
  lowres_df[i,]$dmc_colour <- cluster_dmc[cluster_dmc$cluster==lowres_df[i,]$cluster,]$dmc_colour
  lowres_df[i,]$name <- cluster_dmc[cluster_dmc$cluster==lowres_df[i,]$cluster,]$dmc[[1]]$hex
}
dmc_frame <- tibble()
for(i in 1:k){
  dmc_frame <- rbind(dmc_frame,c(cluster_dmc$dmc[[i]]$dmc,cluster_dmc$dmc[[i]]$name,cluster_dmc$dmc[[i]]$hex))
}
names(dmc_frame)[1]="dmc_colour"
names(dmc_frame)[2]="name"
names(dmc_frame)[3]="col"

new_name <- vector()
for (i in 1:nrow(dmc_frame)){
  new_name[i] <- paste(dmc_frame$name[i],"(",dmc_frame$dmc_colour[i],")")
}
dmc_frame <- dmc_frame %>% mutate(new_name=new_name)

lowres_df[lowres_df$name==background_color,7] <- "B5200"
dmc_frame[dmc_frame$col==background_color,1] <- "B5200"

if(!black_white){
  plot <- lowres_df %>% ggplot(aes(x, y)) + geom_point(aes(col = factor(dmc_colour), 
                                                   shape = factor(dmc_colour))) +coord_fixed()+
    scale_colour_manual(name="1",values=dmc_frame%>%select(dmc_colour,col)%>%deframe,label=dmc_frame%>%select(dmc_colour,new_name)%>%deframe)+
    scale_shape_manual(name="1",values=c(0:10),labels=dmc_frame%>%select(dmc_colour,new_name)%>%deframe)+
    scale_y_reverse()+theme_linedraw()+
    theme(legend.title = element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.position = "left")
     }
else{
  plot<- lowres_df %>% ggplot(aes(x, y)) + geom_point(aes(shape = factor(dmc_colour))) +coord_fixed()+
    scale_colour_manual(name="1",values=dmc_frame%>%select(dmc_colour,col)%>%deframe,label=dmc_frame%>%select(dmc_colour,new_name)%>%deframe)+
    scale_shape_manual(name="1",values=c(0:10),labels=dmc_frame%>%select(dmc_colour,new_name)%>%deframe)+
    scale_y_reverse() + theme_linedraw()+
    theme(legend.title = element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.position = "left")}
  plot

}
```

## Load Images
You can replace "pikachu.jpeg" by your selected picture. The variable 'im' is an object of class 'cimg'.
```{r}
image_file <- "pikachu.jpg"
im <- imager::load.image(image_file)##load the given image to im
plot(im)
```

## Clustering
  We cluster the pixels by its RGB values. The first argument is the image file that needs to be clustered. The second argument is the number of clusters. The output is a data frame of each k and its cluster information. And the nearest DMC thread colour to the original RGB colour is in the data frame as well.
```{r, warning=FALSE, message=FALSE}
set.seed(1005705300)
clusters <- process_image(image_file,2:10) ## Apply clustering to the image with k from 2 to 10
```

## Plot Clusters
  The function scree_plot produces the following scree plots by providing the cluster data frame. In the first graph, it is a tot.withnss vs k plot. We will choose k when the curve gets flat. The second graph is the ratio of k and (k-1) that helps us make the choice.  
  The function colour_strips gives the colour strips of each clusters. The shown colours are the colours of cluster centers.
```{r}
scree_plot(clusters)
colour_strips(clusters)
```  
  
Here, k= 8 might be a good choice.

## Plot the Pattern
At last, we produces the cross-stitches pattern.  
In the make_pattern function, the first argument is the cluster data frame. The second argument is the number of colour we want to have in the pattern. The third argument is the pattern size. The fourth argument is a logic_test. TRUE means black-and-white pattern. FALSE means colorful pattern.The fifth argument is a color name. The selected colour will be replaced by white.  
In this example, although we choose k=8, but we only have 7 colours (including pure white) because of change_resolution in make_pattern.We will not choose background color since it is white. And we want a colorful pattern.
```{r warning=FALSE, message=FALSE}
make_pattern(clusters, 8 ,50,FALSE,NULL)
```
